---
title: "Lab 09: Logistic Regression"
subtitle: "Due: Thu, Apr 15 at 11:59pm ET"
author: "Navya Belavadi"
date: "04-16-2021"
editor_options: 
  chunk_output_type: console
output:
  pdf_document: default
---

```{r set_up, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE,
                      fig.align = "center")
```

## Packages

```{r echo=TRUE}
library(tidyverse)
library(broom)
```

## Data

```{r data}
retirement <- read_csv("data/retirement_data.csv")
```

The 2018 election saw a near-record number of [retirements](https://www.pewresearch.org/fact-tank/2018/04/11/near-record-number-of-house-members-not-seeking-re-election-in-2018/) from Congress. In this lab, you will be working with data from that election.

The variables in this dataset are:

- `stcd`: an ID number for the congressional district. Districts are listed in 
          alphabetic order (e.g., 101 is Alabama's first district).

- `dpres`: the two-party percentage of vote received by Hillary Clinton 
           in the district in 2016. (Third party vote has been eliminated here, 
           this represents Democratic vote/ (Democratic + Republican Vote).)

- `retiring`: a dummy variable indicating whether the representative retired in 
              2018. The includes both representatives who left politics 
              altogether and those who sought higher office, as well as some 
              who resigned early.

- `gopseat`: a dummy variable indicating whether the seat was held by a 
             Republican.

- `age`: he age of the representative at the end of 2018.

# Exercises

## Exercise 1
```{r ex1.1}
new_retirement <- retirement %>%
  group_by(gopseat) %>%
  mutate(dpres = case_when(gopseat == 0 ~ dpres, gopseat == 1 ~ 100-depres), gopseat = ifelse(gopseat == 0, "Democrat", "Republican")) %>%
  filter(!is.na(gopseat) | !is.na(age) | !is.na(retiring))
```

```{r ex1.2}
log_mod <- glm(retiring ~ dpres + factor(gopseat) + age, data = new_retirement, family = "binomial")
tidy(log_mod)
```

## Exercise 2
```{r ex2}
tidy(log_mod) %>%
  mutate(estimate_odds = exp(estimate)) %>%
  select(term, estimate, estimate_odds)
```


## Exercise 3
```{r ex3}
tibble(
  dpres = rep(33:96, times = 2),
  gopseat = factor(rep(c("Democrat", "Republican"), each = 64)),
  age = 59.42) %>%
  augment(log_mod, newdata = .) %>%
  mutate(log_mod, newdata = .) %>%
  ggplot(aes(x = dpres, y = predicted_p, color = gopseat)) +
  geom_line(size = 1.5) +
  theme_minimal(base_size = 14) +
  scale_color_manual(values = c("blue", "red")) +
  labs(title = "Graph of Predicted Probability Values vs. Percentage of Vote", subtitle = "Colored by Political Party", x = "Percentage of Vote Received", y = "Predicted Probability of Retirement", color = "Political Party")
```
It can be observed from this graph that as the percentage of vote received increases for each of the political parties (Democrat and Republican), the predicted probability of retirement decreases. This relationship appears to be stronger for the Republican party than the Democratic party, as the curve of best fit for Republicans decreases at a more rapid rate than that of the Democrats. Therefore, the percentage of vote received increasing had a larger effect on the Republican party than the Democratic party. 

## Exercise 4
```{r ex4}
tibble(
  age = 78,
  gopseat = factor("Democrat"),
  dpres = 70.75) %>%
  augment(log_mod, newdata = .) %>%
    mutate(p = exp(.fitted) / (1 + exp(.fitted))) %>%
    pull(p)
```
According to this logistic model, the probability that Price would retire in 2018 would be 0.0825. 

## Exercise 5
```{r ex5.1}
retirement_probs <- augment(log_mod, newdata = new_retirement) %>%
  mutate(p = exp(.fitted) / (1 + exp(.fitted))) %>%
  arrange(desc(p))
retirement_probs
```

```{r ex5.2}
retirement_probs %>% slice(1)
```

No member of Congress had over a 50% chance of retiring in 2018. The highest probability of retirement belongs to Pennsylvania's 5th district's representative from District 3805, which was at a 40.70% chance -- this representative did end up retiring. 
