---
title: "Lab 09: Logistic Regression"
subtitle: "Due: Thu, Apr 15 at 11:59pm ET"
author: ""
date: ""
editor_options: 
  chunk_output_type: console
output:
  pdf_document: default
---

```{r set_up, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE,
                      fig.align = "center")
```

## Packages

```{r echo=TRUE}
library(tidyverse)
library(broom)
```

## Data

```{r data}
retirement <- read_csv("data/retirement_data.csv")
```

The 2018 election saw a near-record number of [retirements](https://www.pewresearch.org/fact-tank/2018/04/11/near-record-number-of-house-members-not-seeking-re-election-in-2018/) from Congress. In this lab, you will be working with data from that election.

The variables in this dataset are:

- `stcd`: an ID number for the congressional district. Districts are listed in 
          alphabetic order (e.g., 101 is Alabama's first district).

- `dpres`: the two-party percentage of vote received by Hillary Clinton 
           in the district in 2016. (Third party vote has been eliminated here, 
           this represents Democratic vote/ (Democratic + Republican Vote).)

- `retiring`: a dummy variable indicating whether the representative retired in 
              2018. The includes both representatives who left politics 
              altogether and those who sought higher office, as well as some 
              who resigned early.

- `gopseat`: a dummy variable indicating whether the seat was held by a 
             Republican.

- `age`: he age of the representative at the end of 2018.

# Exercises

## Exercise 1



